### 2021-12-16 documentos/kraus.tex
quitando typos?
lots of changes after a lot of time
muevo ecuaciones de dilatacion y semigrupo
etc
Pig

\section{Kraus operators and decoherence} % {{{
% \todo{Carlos hace una version}

@@

We shall first show that the aforementioned Kraus operators indeed produce a PCE. 
Notice that 
% \begin{equation}
$\sigma_\alpha \sigma_\beta \sigma_\alpha =a_{\alpha \beta} \sigma_\beta$,
% \label{}
% \end{equation}
which in turn imply that 
\begin{equation}
\sigma_{\vec \alpha} \sigma_{\vec \beta} \sigma_{\vec \alpha} 
=\san_{\vec \alpha \vec \beta} \sigma_{\vec \beta}.
\label{}
\end{equation}
\cpnote{This is the original $a$, not $H \otimes H$. See how to relate these
two matrices, or define a new $a$ and $A$}
% with $a$ the matrix that appears in \eqref{eq:eigvals-PCE}. 
Next, consider the action of a channel with Kraus representation 
\eqref{eq:kraus:PCE} on an $N$ qubit system:
\begin{equation}
% \begin{align}
\rho & \to 
 \frac{\rho}{2} 
 + \sum_{\vec \beta} \frac{\tau_{\vec \beta}}{2}
\sigma_{\vec \alpha} \sigma_{\vec \beta} \sigma_{\vec \alpha}  
&= 
  \sum_{\vec \beta} 
   \tau_{\vec \beta}\frac{1+\san_{\vec \alpha \vec \beta}}{2}\sigma_{\vec \beta}. 
% 
% 
% \frac{\otimes_k \sigma_{i_k}}{\sqrt{2}} \rho
% \frac{\otimes_k \sigma_{i_k}}{\sqrt{2}} \\
% &=\frac12 \left(\rho + \bigotimes_k \sigma_{i_k} \rho \bigotimes_k \sigma_{i_k} \right)\\
% &=\frac12 \rho
% + \frac12\sum_{j_1, \ldots,j_n} r_{j_1 \ldots j_n} \bigotimes_k  \sigma_{i_k} \sigma_{j_k}  \sigma_{i_k} \\
% &=
% \sum_{j_1, \ldots,j_n} r_{j_1 \ldots j_n} \frac{1+\prod_k a^{i_k}_{j_k}}{2} \bigotimes_k  \sigma_{j_k} . 
\end{equation}
However, since $\san_{\vec \alpha \vec \beta} = \pm 1$, the channel characterized
by Kraus operators  \eref{eq:kraus:PCE} is a PCE. Moreover, One can notice that, 
except for the first row, half of the matrix elements of each row is $+1$ and 
half $-1$, which implies that the aforementioned channel is a PCEG.

@@

\begin{itemize}
\item Construct the Kraus operators, and relate to decoherence, for PCEG
\item Play with non PCEG PCEs
\item Dependiendo de si sale algo bonito, podemos poner algun modelo en
particular para todos los canales (sistema central + ambiente). Seria de buscar
dilataciones sencillas y razonables por lo menos a los canales generadores.
\todo{Davalos lidera la discusión de ese punto, pero no escribe nada}
\end{itemize} 
% }}}

>>>>

\section{Kraus operators and decoherence} % {{{
\cpnote{HAcer esta seccion despues de que david ponga lo de decoherencia, 
para servirla bien}
% \todo{Carlos hace una version}

Furthermore, quantum channels enjoy the well-known \textit{Kraus
representation} (or operator-sum representation) which reads 
\begin{equation}
\mcE[\rho]=\sum_i K_i \rho
K_i^\dagger{},
\end{equation}
with $\sum_i K_i^\dagger{}K_i=\one$ (the trace-preserving condition)~\cite{Kraus1983}.

@@

We shall first show that the aforementioned Kraus operators indeed produce a PCE. 
Notice that 
% \begin{equation}
$\sigma_\alpha \sigma_\beta \sigma_\alpha =a_{\alpha \beta} \sigma_\beta$,
% \label{}
% \end{equation}
which in turn imply that 
\begin{equation}
\sigma_{\vec \alpha} \sigma_{\vec \beta} \sigma_{\vec \alpha} 
=\san_{\vec \alpha \vec \beta} \sigma_{\vec \beta}.
\label{eq:sigma_property}
\end{equation}
\cpnote{This is the original $a$, not $H \otimes H$. See how to relate these
two matrices, or define a new $a$ and $A$}
% with $a$ the matrix that appears in \eqref{eq:eigvals-PCE}. 
Next, consider the action of a channel with Kraus representation 
\eqref{eq:kraus:PCE} on an $N$ qubit system:\janote{acá abajo tenías un $\&$ dentro 
del entorno \texttt{equation} y no compilaba}\janote{había otro en algún lado que quité 
también.}
\begin{equation}
% \begin{align}
\rho 
%& 
\to 
 \frac{\rho}{2} 
 + \sum_{\vec \beta} \frac{\tau_{\vec \beta}}{2}
\sigma_{\vec \alpha} \sigma_{\vec \beta} \sigma_{\vec \alpha}  
= 
  \sum_{\vec \beta} 
   \tau_{\vec \beta}\frac{1+\san_{\vec \alpha \vec \beta}}{2}\sigma_{\vec \beta}. 
% 
% 
% \frac{\otimes_k \sigma_{i_k}}{\sqrt{2}} \rho
% \frac{\otimes_k \sigma_{i_k}}{\sqrt{2}} \\
% &=\frac12 \left(\rho + \bigotimes_k \sigma_{i_k} \rho \bigotimes_k \sigma_{i_k} \right)\\
% &=\frac12 \rho
% + \frac12\sum_{j_1, \ldots,j_n} r_{j_1 \ldots j_n} \bigotimes_k  \sigma_{i_k} \sigma_{j_k}  \sigma_{i_k} \\
% &=
% \sum_{j_1, \ldots,j_n} r_{j_1 \ldots j_n} \frac{1+\prod_k a^{i_k}_{j_k}}{2} \bigotimes_k  \sigma_{j_k} . 
\end{equation}
However, since $\san_{\vec \alpha \vec \beta} = \pm 1$, the channel characterized
by Kraus operators  \eref{eq:kraus:PCE} is a PCE. Moreover, One can notice that, 
except for the first row, half of the matrix elements of each row is $+1$ and 
half $-1$, which implies that the aforementioned channel is a PCEG.

@@

\begin{itemize}
\item Relate to decoherence, for PCEG
\item Play with non PCEG PCEs
\item Dependiendo de si sale algo bonito, podemos poner algun modelo en
particular para todos los canales (sistema central + ambiente). Seria de buscar
dilataciones sencillas y razonables por lo menos a los canales generadores.
\todo{Davalos lidera la discusión de ese punto, pero no escribe nada}
\end{itemize}

\begin{equation}
\mcL_{\vec \alpha}[\Delta]:= \frac{\gamma}{2}\left( \sigma_{\vec\alpha} \Delta \sigma_{\vec\alpha} -\Delta \right).
\label{eq:Nqubit_semigroup}
\end{equation}

\begin{align}
U_{\vec\alpha}\ket{\psi} \ket{0}:=\frac{1}{\sqrt{2}} \left( \ket{\psi} \ket{0}+\sigma_{\vec\alpha} \ket{\psi}\ket{1} \right)\\
U_{\vec\alpha}\ket{\psi} \ket{1}:=\frac{1}{\sqrt{2}} \left( \ket{\psi} \ket{0}-\sigma_{\vec\alpha} \ket{\psi}\ket{1} \right)
\label{eq:Nqubit_stinespring}
\end{align}
% }}}

### 2021-12-16 documentos/pce.tex
Pig

\section{PCE Maps} 
To introduce the family of Pauli component erasing maps we start with 
review basic concepts of quantum channels that will allow us to introduce some 
notation. 
% In this section we introduce the definition of Pauli component erasing maps,
% but first we go through several standard definitions needed to discuss them.
\subsection{Quantum channels} % {{{

@@

The construction of quantum channels includes basically three ingredients:
linearity, trace preservation and complete positivity. Linearity is needed to
map every convex combination of density matrices into a convex combination of the
evolution of such density matrices. The
trace preserving property is required for the process $\mcE$ to be
deterministic, and reads $\tr\mcE[\rho]=\tr \rho =1$. 
The
complete
positivity condition is needed to preserve positive semidefiniteness and handle
the non-local feature of quantum theory. 
A linear map $\mcE$ is positive if it maps density operators to density operators, \ie
if $\mcE[\rho]\geq 0$ for all density matrices $\rho$. 
% 
% $\mcE[\Delta]\geq 0 \ \ \forall \Delta \in \mcB(\hilbert)$ hermitian with $\Delta\geq 0$\cpnote{Francois, de tu nota, la $\Delta$ ya es hermitica, pues 
% decimos que es una matriz de densidad} \ddnote{Aqui hablamos de matrices
% positivas definidas en general, estuve revisando en varios libros, y parece que
% lo mas estandar es suponer que si se dice que es positiva (o negativa) definida
% o semi-definida, automaticamente se dice que es hermítica. Sin embargo hay
% definiciones mas generales que involucran matrices no hermiticas, esas eran las
% que yo tenía en mente (pero parece que no son tan conocidas). Opto por
% mencionar que $\Delta$ es hermítico.}. 
%
On the other hand, if one extends a positive map to include an ancilla, the resulting
map is not always positive. If, for all dimensions of the ancilla, such extension
results in a positive map, we say that the original map is completely
positive~\cite{}.  Quantum channels are required to be completely positive
so as to allow the proper evolution of potentially entangled states with an 
ancilla. 
% In the other hand, a map is completely positive if extending the 
%  $\id_k \otimes \mcE$ is a positive map for $k=1,2,\dots$, where $\id_k$ is the identity channel over $\mcB(\hilbert_k)$. This ensures that every positive semidefinite operator of any
% extension of the system (where the quantum channel is now a local operation) is
% mapped to a positive semidefinite operator. The trace preserving is automatically
% fulfilled for all extensions if $\mcE$ fulfills it. Therefore, any density matrix of a system
% where $\mcE$ acts locally, is sent to a valid density matrix, this is the
% substance of complete positivity.
% \flnote{No creo que lo de Bell state sea conveniente. Mejor decir algo como: E
% is totally positive if trivially extending E to the tensor product of H with an
% arbitrary dimensional ancilla space yields a positive operation.
% }\ddnote{ejecutado} 
Concluding, quantum channels are linear completely positive trace preserving
maps.

Quantum channels can be characterized by means of the Choi-\jami{}
isomorphism~\cite{choi,jamil}. This is, given a linear operation
$\mcE:\mcB(\hilbert)\to \mcB(\hilbert)$, the so-called Choi matrix of it is
$\choi=\left(\id \otimes \mcE\right)[\proj{\Omega}{\Omega}]$ (also called dynamical matrix), where
$\ket{\Omega}=1/\dim(\hilbert)\sum_i^{\dim(\hilbert)}\ket{i}\ket{i}$ is the
Bell state between two copies of \hilbert. The map \mcE is completely positive if and only
\choi is positive-semidefinite.
%
%
Furthermore, quantum channels enjoy the well-known \textit{Kraus representation} (or
operator-sum representation) which reads 
\begin{equation}
\mcE[\rho]=\sum_i K_i \rho
K_i^\dagger{},
\end{equation}
with $\sum_i K_i^\dagger{}K_i=\one$ (the trace-preserving condition). In fact, a linear operation is completely positive if and only
if it has a Kraus representation~\cite{Kraus1983}. The minimum number of Kraus operators $K_i$ needed to represent the channel is called \textit{Kraus rank}. 
%

@@

\flnote{
probablemente convenga mostrar la ecuaci'on desplegada. Asimismo se debe decir que 
$\sum_i K_i^\dagger K_i = 1$ para la trace-preserving property. 
Creo qie ``hermiticity preserving'' es en parte consecuencia de la definición, pero esto no es muy importante
}\ddnote{ejecutado, hacia falta edicion, sin embargo lo de hermiticity preserving es para hacer incapie en por que podemos trabajar en una representacion totalmente real de los canales.}

  \ddnote{[Choi representation, Kraus operators and Stinespring dilation]}

@@

\subsection{One qubit \pce{} maps} % {{{

@@

\subsection{$\mathbf{N}$ qubits Pauli component erasing maps} % {{{

@@

As mentioned above, \pce{} maps nullify Pauli strings, thus, consider the linear operations acting in the generalized Bloch vector $\vec r_{\vec \alpha}$,
\begin{equation}
	\paulicomponents \mapsto \taus \paulicomponents.
% 	\taus = 0,1, \hspace{1.5cm}	\tau_{0,\ldots,0}=1,
\label{eq:definition:Pauli}
\end{equation}
with $\tau_{\vec \alpha}$ either $1$ or $0$.
Similar to the qubit case, not every operation with this form is a \cptp{} map. For the operation to be trace preserving, it is easy to prove that $\tau_{\vec 0}=1$. On the other hand, constructing and evaluating the conditions for complete positivity is involved and is the main problem addressed in this paper. Moreover, these operations are by definition diagonal in the $\left\{ 2^{-N/2}\sigma_{\vec
\alpha} \right\}$ basis.

>>>>

\section{PCE Maps} 
To introduce the family of Pauli component erasing maps we start with 
review basic concepts of quantum channels that will allow us to introduce some 
notation. 
% In this section we introduce the definition of Pauli component erasing maps,
% but first we go through several standard definitions needed to discuss them.
\subsection{Quantum channels} % {{{
\cpnote{Cambios importantes, qeu David revise}

@@

The construction of quantum channels includes basically three ingredients:
linearity, trace preservation and complete positivity. Linearity is needed to
map every convex combination of density matrices into a convex combination of the
evolution of such density matrices. The
trace preserving property is required for the process $\mcE$ to be
deterministic, and reads $\tr\mcE[\rho]=\tr \rho =1$. 
Often, this condition is easy to test.
The
complete
positivity condition is needed to preserve positive semidefiniteness and handle
the non-local feature of quantum theory. 
A linear map $\mcE$ is positive if it maps density operators to density operators, \ie
if $\mcE[\rho]\geq 0$ for all density matrices $\rho$. 
% 
% $\mcE[\Delta]\geq 0 \ \ \forall \Delta \in \mcB(\hilbert)$ hermitian with $\Delta\geq 0$\cpnote{Francois, de tu nota, la $\Delta$ ya es hermitica, pues 
% decimos que es una matriz de densidad} \ddnote{Aqui hablamos de matrices
% positivas definidas en general, estuve revisando en varios libros, y parece que
% lo mas estandar es suponer que si se dice que es positiva (o negativa) definida
% o semi-definida, automaticamente se dice que es hermítica. Sin embargo hay
% definiciones mas generales que involucran matrices no hermiticas, esas eran las
% que yo tenía en mente (pero parece que no son tan conocidas). Opto por
% mencionar que $\Delta$ es hermítico.}. 
%
On the other hand, if one extends a positive map to include an ancilla, the resulting
map is not always positive. If, for all dimensions of the ancilla, such extension
results in a positive map, we say that the original map is completely
positive~\cite{}.  Quantum channels are required to be completely positive
so as to allow the proper evolution of potentially entangled states with an 
ancilla. 
% In the other hand, a map is completely positive if extending the 
%  $\id_k \otimes \mcE$ is a positive map for $k=1,2,\dots$, where $\id_k$ is the identity channel over $\mcB(\hilbert_k)$. This ensures that every positive semidefinite operator of any
% extension of the system (where the quantum channel is now a local operation) is
% mapped to a positive semidefinite operator. The trace preserving is automatically
% fulfilled for all extensions if $\mcE$ fulfills it. Therefore, any density matrix of a system
% where $\mcE$ acts locally, is sent to a valid density matrix, this is the
% substance of complete positivity.
% \flnote{No creo que lo de Bell state sea conveniente. Mejor decir algo como: E
% is totally positive if trivially extending E to the tensor product of H with an
% arbitrary dimensional ancilla space yields a positive operation.
% }\ddnote{ejecutado} 
% Concluding, quantum channels are linear completely positive trace preserving
% maps.

A simple algorithm to test the complete positivity of a quantum channel 
was developed by 
\jami{}~\cite{jamil} and Choi~\cite{choi}.
One first exploits an isomorphism that maps a channel $\mcE$ to the state
$\choi=\left(\id \otimes \mcE\right)[\proj{\Omega}{\Omega}]$, where
$\ket{\Omega}=1/\dim(\hilbert)\sum_i^{\dim(\hilbert)}\ket{i}\ket{i}$ is a 
maximally entangled state between the original system and an ancilla.
Remarkably, the map \mcE is completely positive
if and only \choi (also called dynamical matrix) is positive-semidefinite.

%  the
% Bell state between two copies of \hilbert.

% Quantum channels can be characterized by means of the Choi-\jami{}
% isomorphism~\cite{choi,jamil}. This is, given a linear operation
% $\mcE:\mcB(\hilbert)\to \mcB(\hilbert)$, the so-called Choi matrix of it is
% $\choi=\left(\id \otimes \mcE\right)[\proj{\Omega}{\Omega}]$ (also called
% dynamical matrix), where
% $\ket{\Omega}=1/\dim(\hilbert)\sum_i^{\dim(\hilbert)}\ket{i}\ket{i}$ is the
% Bell state between two copies of \hilbert. The map \mcE is completely positive
% if and only \choi is positive-semidefinite.
%
%

% In fact,
% a linear operation is completely positive if and only if it has a Kraus
% representation~\cite{Kraus1983}.
% The minimum number of Kraus operators $K_i$
% needed to represent the channel is called \textit{Kraus rank}. 
%

@@

% \flnote{
% probablemente convenga mostrar la ecuaci'on desplegada. Asimismo se debe decir que 
% $\sum_i K_i^\dagger K_i = 1$ para la trace-preserving property. 
% Creo qie ``hermiticity preserving'' es en parte consecuencia de la definición, pero esto no es muy importante
% }\ddnote{ejecutado, hacia falta edicion, sin embargo lo de hermiticity preserving es para hacer incapie en por que podemos trabajar en una representacion totalmente real de los canales.}
% 
%   \ddnote{[Choi representation, Kraus operators and Stinespring dilation]}

@@

\subsection{Pauli component erasing maps} % {{{

@@

% \subsection{$\mathbf{N}$ qubits Pauli component erasing maps} % {{{

@@

As mentioned above, \pce{} maps nullify Pauli strings, thus, consider the linear operations acting in the generalized Bloch vector $\vec r_{\vec \alpha}$,
\begin{equation}
	\paulicomponents \mapsto \taus \paulicomponents.
% 	\taus = 0,1, \hspace{1.5cm}	\tau_{0,\ldots,0}=1,
\label{eq:definition:Pauli}
\end{equation}
with $\tau_{\vec \alpha}$ either $1$ or $0$.
\cpnote{Revisar aca lo que sigue}
Similar to the qubit case, not every operation with this form is a completely
positive trace perserving map. For the operation to be trace preserving, it is
easy to prove that $\tau_{\vec 0}=1$. On the other hand, constructing and
evaluating the conditions for complete positivity is involved and is the main
problem addressed in this paper. Moreover, these operations are by definition
diagonal in the $\left\{ 2^{-N/2}\sigma_{\vec
\alpha} \right\}$ basis.
